Calendar-based graphics for visualizing people's daily schedules

### 2019-05-14 main.Rmd
Minor wording tweaks

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \@ref(fig:ped-map)). The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \@ref(fig:facet-time)) provides a better glimpse of the daily and sub-daily pedestrian patterns. Faceting generates small multiples [@tufte1983visual] or trellis displays [@becker_visual_1996], where different subsets of the same data are displayed based on conditioning variables, allowing for comparison across subsets. A series of graphics can be further grouped together to create multiple yet different data views, that forms ensemble graphics [@unwin_ensemble_2018].

However, the conventional displays of time series data conceal patterns relative to special events (such as public holidays and recurring cultural/sport events), which may be worth noting to viewers.

@@

The grid position for any day in the month is given by

\begin{equation}
  \begin{aligned}
  i &= \lceil (g \text{ mod } 35) / 7\rceil, \\
  j &= g \text{ mod } 7.
  \end{aligned}
  (\#eq:grid)
\end{equation}

@@

Then the final points for making the calendar line plots of the pedestrian sensor data is given by:

\begin{equation}
  \begin{aligned}
  x &= j + (n - 1) \times 7 + (n - 1) \times b + h, \\
  y &= i - (m - 1) \times 5 - (m - 1) \times b + c.
  \end{aligned}
  (\#eq:final)
\end{equation}

@@

```
frame_calendar(data, x, y, date, calendar = "monthly", dir = "h", 
  week_start = 1, nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed", 
  width = 0.95, height = 0.95, margin = NULL)
```

@@

The calendar-based visualization provides data plots in the familiar format of an everyday tool. Patterns on special events and public holidays for the region are more visible to the viewer.

The calendar layout will be useful for studying consumer trends and human behavior. It will be less useful for physical patterns like climate, which are not typically affected by human activity. The layout does not replace traditional displays, but serves to complement to further tease out structure in temporal data. Analysts would still be advised to plot overall summaries and deviations in order to study general trends.

>>>>

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped], using installed sensors that count pedestrians every hour across the inner-city area. We use data from 43 sensors until the end of 2016 (Figure \@ref(fig:ped-map)). The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \@ref(fig:facet-time)) provides a better glimpse of the daily and sub-daily pedestrian patterns. Faceting generates small multiples [@tufte1983visual] or trellis displays [@becker_visual_1996], where different subsets of the same data are displayed based on conditioning variables, allowing for comparison across subsets. A series of graphics can be further grouped together to create multiple yet different data views, that forms ensemble graphics [@unwin_ensemble_2018].

However, the conventional displays of time series data conceal patterns relative to special events (such as public holidays and recurring cultural/sport events), which may be of interest to viewers.

@@

The grid position for any day in the month is given by
\begin{equation}
  \begin{aligned}
  i &= \lceil (g \text{ mod } 35) / 7\rceil, \\
  j &= g \text{ mod } 7.
  \end{aligned}
  (\#eq:grid)
\end{equation}

@@

Then the final points for making the calendar line plots of the pedestrian sensor data is given by:
\begin{equation}
  \begin{aligned}
  x &= j + (n - 1) \times 7 + (n - 1) \times b + h, \\
  y &= i - (m - 1) \times 5 - (m - 1) \times b + c.
  \end{aligned}
  (\#eq:final)
\end{equation}

@@

```r
frame_calendar(data, x, y, date, calendar = "monthly", dir = "h",
  week_start = 1, nrow = NULL, ncol = NULL, polar = FALSE, scale = "fixed",
  width = 0.95, height = 0.95, margin = NULL)
```

@@

The calendar-based visualization provides data plots in the familiar format of an everyday tool. Patterns relating to special events and public holidays for the region are more visible to the viewer.

The calendar layout will be useful for studying consumer trends and human behavior. It will be less useful for physical patterns like climate, which are not typically affected by human activity. The layout does not replace traditional displays, but serves to complement them and to further tease out structure in temporal data. Analysts would still be advised to plot overall summaries and deviations in order to study general trends.
